Automated topic naming to support cross-project analysis of software maintenance activities

### 2011-03-25 msr/msr2011.tex
intro and previous work

\newcommand{\shrink}{}
\section{Introduction}

@@

%One challenge for software maintenance is to identify the reasons
%behind software development activities. 
%Identifying these reasons provides insight into \emph{why} certain
%changes were made. 
%migod
A key problem for practising software maintainers is gaining an
understanding of \emph{why} a system has evolved the way it has. 
Looking back on streams of artifacts scattered across different
repositories, inferring what activities were performed, when, and for
what reasons, is hard to do without expert advice from the developers
involved. 
%new stuff
% edit this. awkward
This is different than \emph{how}, because the behaviour of change itself is the
\emph{how}.
In this work we seek to provide automated in labelling development topics extracted from commit logs.

Topic modeling is a machine learning technique which creates
multinomial distributions over words from a text corpus. 
The technique infers the hidden structure of a corpus using posterior
inference: the probability of the hidden structure given the data. 
Topic models are useful in software maintenance because they summarize
the key concepts in a corpus, such as source code, commit comments, or
mailing-list messages, by identifying statistically co-occurring words. 
Among other uses, it can quickly give developers an overview of where significant
activity has occurred, and gives managers or maintainers a sense of
project history. 
%A major challenge for topic modeling is to accurately identify the
%summary label for a topic; in the previous case, perhaps
%\emph{``event-handling''} would be appropriate. 
%Current approaches are manual: they rely on human expertise to
%recognize the appropriate label. 
%Furthermore, to date the results of topic modeling are project
%specific, and not generalizable. 
%This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.

%migod
While machine learning techniques can automatically identify clumps of
commonly recurring terms, devising an appropriate summary label for
each clump/topic is harder.  
A given topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}. 
This topic might reasonably be labelled as
``\emph{event handling}'' by a human who understands the domain well,
despite the fact that label does not appear in the word list itself.  
Current approaches to topic labelling rely on manual intervention by
human experts, and also are limited to project-specific topic labels.  
In this paper, we introduce \emph{labelled topic extraction}, an
approach to topic labelling that creates labels automatically that are
project independent.
%is both automatic and creates labels that are project independent.

@@

Concrete applications of topics and \emph{labelled topic extraction}
range from project dashboards to annotating software artifacts such as revisions and bug reports with
NFR-related tags.
Project dashboard~\cite{dashboard} are typically employed by managers and are used to
provide quick summaries of the effort put into a software project. In this case, 
labelled topic extraction would allow managers to track effort
related to NFR topics, such as usability.
These techniques also allow for the annotation of commit comments
 and other software artifacts with NFRs. 
This would enable querying of bug reports and artifacts by relevant NFRs. 
For instance a manager can confirm if their developers were
focused on \emph{usability} by looking for
 \emph{usability}-relevant revisions and bug reports.

@@

\item evaluates three kinds of automatic topic labelling methods:
  semi-unsupervised labelling (word-lists) of topics, supervised labelling of
  topics with a single NFR (machine learning), 
  and supervised labelling of topics with multiple NFRs (multi-label
  machine learning)

@@

Cleland-Huang and her colleagues have investigated on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
Their approach is similar to ours, as they mined keywords from NFR
catalogues.
They differ because they mine requirements documents where as we mine revisions.
%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
In this paper we make cross-project comparisons instead of focusing
on a single project.
%They relied on requirements documents instead of version control
%histories that we use.
%The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project.

@@

Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
They used WordNet for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
Mockus et al. validated their labels with system developers.
Since we study multiple projects, instead of a single project, these
kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).

Another approach is to extract concerns from software repositories.
Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI) to identify commonly occurring concerns for software maintenance. The concerns are given by the user, and LSI used to retrieve them from a corpus. Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.
%Some results were interesting, but their precision was quite low. 
With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence using colour intensity. 
They mine
developer created tags 
%change request tags (such as `milestone 3') and used these 
in order to analyze the evolution of a single product.
% make evolutionary analyses of a single product. 
The presence of a well-maintained set of tags is obviously essential to the success of this technique.

@@

% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
experts read the highest-frequency members of a topic and assign a
label accordingly. 
%XXX Possible to cut this
As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called \emph{`Oracle'} in the MySQL case, since Oracle owns MySQL. 
We do this differently.
%This paper does this differently. 
First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.

@@

\subsubsection{Analysis of the semi-unsupervised labelling} %--- %Based
                                %on the labels, and our manual topic
                                %labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.

>>>>

\newcommand{\shrink}{}

\vspace*{1em}
\section{Introduction}

@@

%One challenge for software maintenance is to identify the reasons
%behind software development activities. 
%Identifying these reasons provides insight into \emph{why} certain
%changes were made. 
%migod
A key problem for practising software maintainers is gaining an
understanding of \emph{why} a system has evolved the way it has. 
This is different than \emph{how} a system has evolved, because the behaviour of change itself is the
\emph{how}.
Looking back on streams of artifacts scattered across different
repositories, inferring what activities were performed, when, and for
what reasons, is hard to do without expert advice from the developers
involved. 
%new stuff
% edit this. awkward
In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.

Topic modeling is a machine learning technique which creates
multinomial distributions of words extracted from a text corpus. 
The technique infers the hidden structure of a corpus using posterior
inference: the probability of the hidden structure given the data. 
Topic models are useful in software maintenance because they summarize
the key concepts in a corpus, such as source code, commit comments, or
mailing-list messages, by identifying statistically co-occurring words. 
Among other uses, it can quickly give developers an overview of where significant
activity has occurred, and gives managers or maintainers a sense of
project history. 
%A major challenge for topic modeling is to accurately identify the
%summary label for a topic; in the previous case, perhaps
%\emph{``event-handling''} would be appropriate. 
%Current approaches are manual: they rely on human expertise to
%recognize the appropriate label. 
%Furthermore, to date the results of topic modeling are project
%specific, and not generalizable. 
%This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.

%migod
While machine learning techniques can automatically identify clumps of
commonly recurring terms, devising an appropriate summary label for
each clump/topic is harder.  
A given topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}. 
This topic might reasonably be labelled as
``\emph{event handling}'' by a developer who understands the domain well,
despite the fact that this label does not appear in the word list itself.  
Current approaches to topic labelling rely on manual intervention by
human experts, and also are limited to project-specific topic labels.  
In this paper, we introduce \emph{labelled topic extraction}, an
approach to topic labelling that creates labels automatically that are
project independent.
%is both automatic and creates labels that are project independent.

@@

Concrete applications of topics and \emph{labelled topic extraction}
range from project dashboards to annotating software artifacts such as revisions and bug reports with
NFR-related tags.
Project dashboard~\cite{dashboard} are typically employed by managers and are used to
provide quick summaries of the effort put into a software project. In this case, 
labelled topic extraction would allow managers to track effort
related to NFR topics, such as portability.
These techniques also allow for the annotation of commit comments
 and other software artifacts with NFRs. 
This would enable querying of bug reports and artifacts by relevant NFRs. 
For instance a manager can confirm if their developers were
focused on \emph{usability} by looking for
 \emph{usability}-relevant revisions and bug reports.

@@

\item evaluates three kinds of automatic topic labelling methods:
  semi-unsupervised labelling  of topics (word-lists), supervised labelling of
  topics with a single NFR (machine learning), 
  and supervised labelling of topics with multiple NFRs (multi-label
  machine learning);

@@

Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
Their approach is similar to ours, as they mined keywords from NFR
catalogues.
They differ because they mine requirements documents where as we mine revisions.
%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
Another difference is that we make cross-project comparisons instead of focusing
on a single project.
%They relied on requirements documents instead of version control
%histories that we use.
%The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project.

@@

Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
They used WordNet for word roots as they felt the synonyms would be
non-specific and cause errors.
%, problems we also encountered. 
Mockus et al. validated their labels with system developers.
Since we study multiple projects, instead of a single project, these
kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).

Another approach is to extract concerns from software repositories.
Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI)
to identify commonly occurring concerns for software maintenance. 
The
concerns are given by the user, and LSI is used to retrieve them from
a corpus. 
Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.
%Some results were interesting, but their precision was quite low. 
With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence
using colour and intensity. 
They mine
developer created tags 
%change request tags (such as `milestone 3') and used these 
in order to analyze the evolution of a single product.
% make evolutionary analyses of a single product. 
The presence of a well-maintained set of tags is obviously essential to the success of this technique.

@@

% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
experts read the highest-frequency members of a topic and assign a
label accordingly. 
%XXX Possible to cut this
As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called \emph{`Oracle'} in the MySQL case, since Oracle owns MySQL. 
We do this differently.
%This paper does this differently. 
First of all, we automate the process of naming the topics. Secondly, we label topics with project-independent terms, in order to permit cross-project comparison.

@@

\subsubsection{Analysis of the semi-unsupervised labelling} %--- %Based
                                %on the labels, and our manual topic
                                %labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labelling process.
As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.
